Observation of S-wave scattering of fermions accross a Feshbach resonance

### 2014-12-30 draft1.tex
Pre-xmas

\documentclass[12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
\usepackage{iopams}  
\usepackage{braket}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cite}

@@

\section{Introduction}
Feshbach resonances are a staple tool in the study of cold Fermi gases. Interatomic interactions in Bose-Einstein condensates are easily detected due to the high density of the cloud. However, the density of Fermi clouds is reduced by a factor of $~10^3$ from that of BECs. This makes it necessary to enhance the scattering cross section to observe any effects of interaction. Feshbach resonances allow one to tune the scattering length by changing the magnetic field. Not ony do they enable detection of interactions, but they make it possible to tune interactions from attractive to repulsive, allowing for the phase transition from the BCS to BEC regime at sufficiently low temperatures. 
\par However, to take advantage of Feshbach resonances as a tool it  is first necessary to have a good undestanding of the relationship between the bias magnetic field and the scattering cross section near a Feshbach resonance for specific species and hyperfine states.  The exact value of the resonance is difficult to calculate analytically and must be determined experimentally. Currently, Feshbach resonances are mapped out by observing loss due to three-body inelastic scattering or by re-thermalization or anisotropic expansion measurements that infer the elastic scattering cross section from collective behavior of the cloud. 
\par We collide two clouds of $^{40}K$ atoms in a mixture of $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$ hyperfine states and directly image the resulting s-wave scattering halo as a function magnetic field strength. This allows us to directly observe the enhancement in scattering without relying on collective behavior. However, even when the scattering is enhanced by the Feshbach resonance, the Fermi cloud is dilute enought that the number of scattered attoms is still very low compared to the number of atoms that pass through the other cloud unscattered. This poses a challenge for imaging both high and low atom numbers in the same shot.
\par In order to optimize the signal to noise for low atom numbers, we increased our imaging time.  However, above saturation absorption imaging at sufficiently long exposure times poses a particular problem - if the atoms have had time to absorb enough photons such that the recoil velocity Doppler shifts them from resonance on the order of the linewidth of the resonance, this recoil-induced detuning is a significant correction to the observed absorption of the probe light. In order to take this correction into account, we simulate the imaging process and use the results to extract the atom number and the scattered fraction from our images.
\par This paper is in two parts. In the first, we study absorption imaging in the presence of significant recoil-induced detuning and show how we use our results to interpret data. In the second, we describe our s-wave scattering experiment and, after correcting our data for recoil-induced detuning, extract a measure of the location of the Feshbach resonance in $^{40}K$.

@@

@@

>>>>

\documentclass[12pt]{iopart}
\usepackage{iopams}  
\usepackage{braket}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cite}

@@

\section{Introduction}

\par In order to optimize the signal to noise for low atom numbers, we increased our imaging time.  However, above saturation absorption imaging at sufficiently long exposure times poses a particular problem - if the atoms have had time to absorb enough photons such that the recoil velocity Doppler shifts them from resonance on the order of the linewidth of the resonance, this recoil-induced detuning is a significant correction to the observed absorption of the probe light. In order to take this correction into account, we simulate the imaging process and use the results to extract the atom number and the scattered fraction from our images.
\par This paper is in two parts. In the first, we study absorption imaging in the presence of significant recoil-induced detuning and show how we use our results to interpret data. In the second, we describe our s-wave scattering experiment and, after correcting our data for recoil-induced detuning, extract a measure of the location of the Feshbach resonance in $^{40}K$.

@@

Feshbach resonances are a staple tool in the study of cold Fermi gases. Interatomic interactions in Bose-Einstein condensates are easily detected due to the high density of the cloud. However, the density of Fermi clouds is reduced by a factor of $~10^3$ from that of BECs. This makes it necessary to enhance the scattering cross section to observe any effects of interaction. Feshbach resonances allow one to tune the scattering length by changing the magnetic field. Not ony do they enable detection of interactions, but they make it possible to tune interactions from attractive to repulsive, allowing for the phase transition from the BCS to BEC regime at sufficiently low temperatures. 
\par However, to take advantage of Feshbach resonances as a tool it  is first necessary to have a good undestanding of the relationship between the bias magnetic field and the scattering cross section near a Feshbach resonance for specific species and hyperfine states.  The exact value of the resonance is difficult to calculate analytically and must be determined experimentally. Currently, Feshbach resonances are mapped out by observing loss due to three-body inelastic scattering or by re-thermalization or anisotropic expansion measurements that infer the elastic scattering cross section from collective behavior of the cloud \cite{Regal03,OHara02,Monroe93}. 
\par We collide two clouds of $^{40}K$ atoms in a mixture of $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$ hyperfine states and directly image the resulting s-wave scattering halo as a function magnetic field strength. This allows us to directly observe the enhancement in scattering without relying on collective behavior. We calculate the fraction of atoms that scattered during the collision, and from this scattering fraction deduce the resonant magnetic field of the resonance.
\par The techniques developed in this experiment for observing Fermion scattering can be extended to engineering higher order partial wave interactions, as has been done for bosons \cite{Williams2012}.

@@

We load our cold Fermi gas of $^{40}K$ atoms into a crossed dipole trap. We then use adiabatic rapid passage to bring the gas into an equal superposition of $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$, the states whose Feshbach resonance we are trying to study. 
\par We then use a large pair of  coils in Helmholtz configuration to bring the magnetic field close to the Feshbach resonance, but not close enough for significant collisions to occur. We hold the atoms there to allow the eddy currents in the large coils to settle, and then use a smaller set of Helmholtz coils to hop the field to the desired value. We take two sets of data: one coming from below the resonance, where we hop the large coils to a lower field value than the resonance and use the small coils to hop up to the desired field, and one coming from below the resonance, where we hop the large coils to a field value above the resonance and then use the small coils to hop the field down to the desired value. This allows us to correct for losses due to molecule formation and three-body recombination \cite{Chin10}.
\par We then use a double pulse sequence \cite{Wu05} of a near resonant 1-d retro-reflected optical lattice to transfer most of the atoms into the first two excited states of the lattice, giving them $\pm 2k_r$ of momentum, where $k_r$ is the recoil momentum of the lattice. Since the initial Fermi gas has a wide momentum spread (in contrast to a BEC, which has a very narrow momentum spread), and the lattice pulsing is a momentum dependent process, not all the atoms are successfully transferred into the 1st excited band of the lattice. We optimized our pulse times to minimize the number of atoms that remain in the lowest, zero momentum, band of the lattice, but were not able to eliminate them completely. The optimized pulse times are ??FILL IN??
\par We release the atoms from the trap and allow ??FILL IN TIME?? for the two opposite momentum states within the cloud to pass through each other, scattering on the way. For the data taken coming from below the Feshbach resonance, we then simply ramp down the field and image the atoms. For the data taken coming from above the Feshbach resonance, we then ramp the field back up through the resonance to recover any molecules that were created when the field was ramped from the attractive to the repulsive side of the resonance, and then quickly ramp the field back down and image the atoms. We use an imaging pulse of 40$\mu s$.
